Evaluating recommender behavior for new users

### 2014-05-06 paper/main.tex
first stab at descriptors, abstract, etc.
cleaned up plots a little

@@

%TODO: update these things.
% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%TODO: update these things.
\terms{Theory}

%TODO: update these things.
\keywords{ACM proceedings, \LaTeX, text tagging}

@@

\subsection{Prediction accuracy}
  % Prediction accuracy
\begin{figure*}[ht!]
  \centering
  \includegraphics[width=0.9\linewidth]{../lenskit/output/ekstrandTuned20/accuracy.pdf}
  \caption{Left: RMSE, Right: nDCG}
  \label{fig:rmse}
  \label{fig:ndcg}
\end{figure*}
    % RMSE plot
    % point out error bars etc.
    % explain / interpret plot
  \paragraph{RMSE}
  Figure \ref{fig:rmse} shows how accurate our algorithms are as measured by RMSE.
  User-User, and SVD both behaved essentially as expected.
  SVD shows no improvement with only one rating, and User-User, actually appears to get worse with the first rating, but other than that, these algorithms outperform the item baseline.
  Interestingly, the User-Item baseline outperforms both algorithms until around 8 ratings, at which point both algorithms start to perform better.
  Given the size of of the differences between UserItemBaseline, UserUser, and SVD we will conclude that all three are at about the same accuracy for new users.
  That such a simple algorithm can outperform for such small values goes to show that recommending for users with very few ratings is a hard problem.

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1.05\linewidth]{../lenskit/output/ekstrandTuned20/TopNPrecision.pdf}
  \caption{Left: Precision@20, Center: MAP@20, Right: Fallout@20}
  \label{fig:map}
\end{figure*}
\paragraph{MAP@20, Precision@20, Fallout@20}
  Figure \ref{fig:map} shows MAP@20, precision@20, and fallout@20.
  All three metrics show essentially the same trend, User-User gets the lowest score, the baseline get the highest score.
  On both MAP@20 and precision@20 svd gets a higher score than Item-Item throughout, performing almost as highly as the baseline.
  On fallout@20 Item-Item and SVD perform about as well as each other, with SVD showing a very slight lead.
  We also tested recall and mean reciprocal rank metrics, but found them to exhibit the same trend.

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1.05\linewidth]{../lenskit/output/ekstrandTuned20/rmse_20.pdf}
  \caption{Left: SeenItems@20, Center: MeanRating@20, Right: RMSE@20}
  \label{fig:topN.rmse}
\end{figure*}
  \paragraph{SeenItems@20}
  The first graph in figure \ref{fig:topN.rmse} shows the average number of items in the top 20 recommendations for each algorithm that were also in the test set.
  The average number of test set items in recommendations shows a similar trend to the MAP, precision, and recall metrics, user-user getting by far the fewest seen items, and the baseline getting the most.
  As User-User gets less than one seen item on average we found that MeanRating@20 and RMSE@20 metrics were not appropriate for UserUser.
  This was observed in particular when testing our results for sensitivity to recommendation list size.
  We found that different recommendation lists led UserUser (and only UserUser) to have very different results
  Because of this issue we will not include User-User in our next evaluation and consider the next two metrics inconclusive for UserUser
  That said, we consider the small number of seen items itself a bad result for UserUser.
  Put simply, if we cannot reliably estimate if the recommendations are reasonable, how can we expect a user to do so?

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=1.05\linewidth]{../lenskit/output/ekstrandTuned20/popdiv.pdf}
  \caption{Left:averagePopularity@20, Center AILS@20, Right Spread@20}
  \label{fig:pop}
\end{figure*}

>>>>

%TODO: cleanup plots.

@@

%TODO: update these things.
% A category with the (minimum) three required fields
\category{H.3.3}{nformation Search and Retrieval}{Information filtering}

%TODO: update these things.
\terms{Algorithms, Measurement}

%TODO: update these things.
\keywords{Recommender Systems, Evaluation, Profile Size, New User Experience, New User Problem, User Cold Start}

@@

\subsection{Prediction accuracy}
  % Prediction accuracy
\begin{figure*}[ht!]
  \centering
  \includegraphics[width=0.666\linewidth]{../lenskit/output/ekstrandTuned20/accuracy.pdf}
  \caption{Left: RMSE, Right: nDCG}
  \label{fig:rmse}
  \label{fig:ndcg}
\end{figure*}
    % RMSE plot
    % point out error bars etc.
    % explain / interpret plot
  \paragraph{RMSE}
  Figure \ref{fig:rmse} shows how accurate our algorithms are as measured by RMSE.
  User-User, and SVD both behaved essentially as expected.
  SVD shows no improvement with only one rating, and User-User, actually appears to get worse with the first rating, but other than that, these algorithms outperform the item baseline.
  Interestingly, the User-Item baseline outperforms both algorithms until around 8 ratings, at which point both algorithms start to perform better.
  Given the size of of the differences between UserItemBaseline, UserUser, and SVD we will conclude that all three are at about the same accuracy for new users.
  That such a simple algorithm can outperform for such small values goes to show that recommending for users with very few ratings is a hard problem.

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\linewidth]{../lenskit/output/ekstrandTuned20/TopNPrecision.pdf}
  \caption{Left: Precision@20, Center: MAP@20, Right: Fallout@20}
  \label{fig:map}
\end{figure*}
\paragraph{MAP@20, Precision@20, Fallout@20}
  Figure \ref{fig:map} shows MAP@20, precision@20, and fallout@20.
  All three metrics show essentially the same trend, User-User gets the lowest score, the baseline get the highest score.
  On both MAP@20 and precision@20 svd gets a higher score than Item-Item throughout, performing almost as highly as the baseline.
  On fallout@20 Item-Item and SVD perform about as well as each other, with SVD showing a very slight lead.
  We also tested recall and mean reciprocal rank metrics, but found them to exhibit the same trend.

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\linewidth]{../lenskit/output/ekstrandTuned20/rmse_20.pdf}
  \caption{Left: SeenItems@20, Center: MeanRating@20, Right: RMSE@20}
  \label{fig:topN.rmse}
\end{figure*}
  \paragraph{SeenItems@20}
  The first graph in figure \ref{fig:topN.rmse} shows the average number of items in the top 20 recommendations for each algorithm that were also in the test set.
  The average number of test set items in recommendations shows a similar trend to the MAP, precision, and recall metrics, user-user getting by far the fewest seen items, and the baseline getting the most.
  As User-User gets less than one seen item on average we found that MeanRating@20 and RMSE@20 metrics were not appropriate for UserUser.
  This was observed in particular when testing our results for sensitivity to recommendation list size.
  We found that different recommendation lists led UserUser (and only UserUser) to have very different results
  Because of this issue we will not include User-User in our next evaluation and consider the next two metrics inconclusive for UserUser
  That said, we consider the small number of seen items itself a bad result for UserUser.
  Put simply, if we cannot reliably estimate if the recommendations are reasonable, how can we expect a user to do so?

@@

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\linewidth]{../lenskit/output/ekstrandTuned20/popdiv.pdf}
  \caption{Left:averagePopularity@20, Center AILS@20, Right Spread@20}
  \label{fig:pop}
\end{figure*}
